MSHS COVID-19 Vaccination Summary

### 2022-07-18 MSH Vaccines Administered Hybrid Data Model.Rmd
Create new markdown file for rendering vaccines administered dashboard using hybrid data approach.

```{r Set directory, import reference data, and repository, include = FALSE}
# # Set working directory and select raw data ----------------------------
# rm(list = ls())
# 
# # Determine path for working directory
# if ("Presidents" %in% list.files("J://")) {
#   user_directory <- paste0("J:/Presidents/HSPI-PM/",
#                            "Operations Analytics and Optimization/Projects/",
#                            "System Operations/COVID Vaccine")
# } else {
#   user_directory <- paste0("J:/deans/Presidents/HSPI-PM/",
#                            "Operations Analytics and Optimization/Projects/",
#                            "System Operations/COVID Vaccine")
# }

@@

# # Determine whether or not to update an existing repo
# update_repo <- FALSE

# Import reference data for site and pod mappings
dept_mappings <- read_excel(paste0(
  user_directory,
  "/Weekly Dashboard Doses Administered",
  "/MSHS COVID-19 Vaccine Department Mappings 2022-04-11.xlsx"))

dept_mappings <- dept_mappings %>%
  select(-`Site (Historical)`, -Notes)

@@

# Vaccine types
vax_types <- c("Adult", "Peds")

# Dept grouper
dept_group <- c("Hospital POD", "MSHS Practice")
```

```{r Import existing repo and append as needed}
# Import schedule repository and determine last date it was updated
sched_repo_file <- choose.files(default = paste0(user_directory,
                                                 "/R_Sched_AM_Repo/*.*"))

sched_repo <- readRDS(sched_repo_file)
sched_repo_update_date <- date(file.info(sched_repo_file)$ctime)

@@

  sched_repo <- sched_repo

@@

```{r Format and process schedule data}
# Format and analyze schedule to date for dashboards ---------------------------
sched_to_date <- sched_repo

sched_to_date <- sched_to_date %>%
  # rename(IsPtEmployee = `Is the patient a Mount Sinai employee`) %>%
  mutate(
    # Update timezones to EST if imported as UTC
    `Appt Time` = with_tz(`Appt Time`, tzone = "EST"),
    # Determine whether appointment is for dose 1, 2, or 3 based on scheduled visit type
    Dose = ifelse(str_detect(Type, "DOSE 1"), "Dose 1",
                  ifelse(str_detect(Type, "DOSE 2"), "Dose 2",
                         ifelse(str_detect(Type, "(DOSE 3)|(BOOSTER)"),
                                "Dose 3/Boosters", NA))),
    # Determine appointment date, year, month, etc.
    ApptDate = date(Date),
    ApptYear = year(Date),
    ApptMonth = month(Date),
    ApptDay = day(Date),
    ApptWeek = week(Date),
    # Clean up department if it is duplicated in that column
    Department = ifelse(str_detect(Department, ","),
                        substr(Department, 1,
                               str_locate(Department, ",") - 1),
                        Department),
    # Group arrived, completed, and checked out appts as arrived
    Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
                    "Arr", `Appt Status`),
    # Update appointment status: classify any appts from prior days that are
    # still in scheduled as No Shows
    Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", 
                     # Correct any appts erroneously marked as arrived early
                     ifelse(ApptDate >= today & Status == "Arr", "Sch",
                            Status)),
    # # Determine whether the vaccine is an adult or pediatric dose
    # VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD",
    #                                    "1468 MADISON HOSP PEDS GEN") |
    #                    `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
    #                                               "DOSE 2 PEDIATRIC [1324685]") |
    #                    Department %in% peds_school_practice_dept,
    #                  "Peds", "Adult"),
    # # Determine manufacturer based on immunization field and vaccine type fields
    # Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
    #   ifelse(Status2 != "Arr", NA,
    #          # Any blank immunizations and any peds assumed to be Pfizer
    #          ifelse(is.na(Immunizations) |
    #                   VaxType %in% c("Peds"), "Pfizer",
    #                 # Any immunizations with Moderna in text classified as Moderna
    #                 ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
    #                        # Any visiting docs or Johnson and Johnson
    #                        # immunizations classified as J&J
    #                        ifelse(str_detect(Department, "VISITING DOCS") |
    #                                 str_detect(Immunizations,
    #                                            "Johnson and Johnson"),
    #                               "J&J", "Pfizer")))),
    # Determine week number of year
    WeekNum = format(ApptDate, "%U"),
    # Determine appointment day of week
    DOW = weekdays(ApptDate)
    )

# Crosswalk sites and department grouper
sched_to_date <- left_join(sched_to_date, dept_mappings,
                           by = c("Department" = "Department"))

sched_to_date <- sched_to_date %>%
  rename(Site = `New Site`)

# Determine if there are any missing departments and notify user
unmapped_dept <- sched_to_date %>%
  filter(is.na(Site)) %>%
  select(Department) %>%
  distinct()

if(nrow(unmapped_dept) > 0) {
  stop(paste0("Check department mappings. Missing departments include ",
             paste(as.vector(unmapped_dept$Department), collapse = ", "),
             "."))
}

# Determine vaccine type (adult vs peds) administered based on department, provider, and age, as appropriate
sched_to_date <- sched_to_date %>%
  mutate(
    # Create a column for age grouper
    Age_Grouper = ifelse(str_detect(Age, "\\ y.o."), "years",
                        ifelse(str_detect(Age, "\\ m.o."), "months",
                               ifelse(str_detect(Age, "\\ days"), "days",
                                      "unknown"))),
    # Convert string with age to numeric
    Age_Numeric = 
      ifelse(Age_Grouper == "years",
                    as.numeric(str_extract(Age,
                                 # pattern = "(?<=Deceased\\ \\().+(?=\\ y.o.)")),
                                 paste("(?<=Deceased\\ \\()[0-9]+(?=\\ y.o.)",
                                       "[0-9]+(?=\\ y.o.)",
                                       sep = "|"))),
                    ifelse(Age_Grouper == "months",
                           as.numeric(str_extract(Age, "[0-9]+(?=\\ m.o.)")) / 12,
                           ifelse(Age_Grouper == "days",
                                  as.numeric(
                                    str_extract(Age, "[0-9]+(?=\\ days)")) / 365,
                                  ifelse(Age_Grouper == "unknown", 10000,
                                         NA_integer_)))),
    # Determine vaccine type based on department, provider/resource, and age if necessary
    VaxType = ifelse(
      `Vaccine Age Group` %in% "Peds" |
        # MSBI Peds POD
        `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
                                       "DOSE 2 PEDIATRIC [1324685]") |
        # Practics that administer adult and pediatric vaccines
        (`Vaccine Age Group` %in% "Both" &
              `Setting Grouper` %in% c("School Based Practice", "MSHS Practice") &
                 Age_Numeric < 12),
      "Peds", "Adult"),
    # Determine manufacturer based on immunization field and vaccine type fields
    Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
      ifelse(Status2 != "Arr", NA,
             # Any blank immunizations and any peds assumed to be Pfizer
             ifelse(is.na(Immunizations) |
                      VaxType %in% c("Peds"), "Pfizer",
                    # Any immunizations with Moderna in text classified as Moderna
                    ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
                           # Any visiting docs or Johnson and Johnson
                           # immunizations classified as J&J
                           ifelse(str_detect(Department, "VISITING DOCS") |
                                    str_detect(Immunizations,
                                               "Johnson and Johnson"),
                                  "J&J", "Pfizer")))),
    Site =
      # Roll up any Moderna administered at MSB for YWCA to MSH
      ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH",
             # Manually classify school practices as peds or adolescent for schools administering both
             ifelse(`Setting Grouper` %in% "School Based Practice" &
                      VaxType %in% "Peds", "Peds School Based Practice",
                    ifelse(`Setting Grouper` %in% "School Based Practice" &
                             VaxType %in% "Adult",
                           "Adolescent School Based Practice",
                           Site))),
    # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
    Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c("Dose 2"), "Dose 1", Dose),
    `Setting Grouper` = ifelse(`Setting Grouper` %in% "School Based Practice",
                               "MSHS Practice", `Setting Grouper`))

vax_admin_start_date <- min(sched_to_date$ApptDate[
  which(sched_to_date$Status2 %in% "Arr")])

vax_admin_end_date <- max(sched_to_date$ApptDate[
  which(sched_to_date$Status2 %in% "Arr")])

@@

## {.tabset}
### All Patient Settings
```{r Create kable of all doses adminsitered to date}
# Summarize data for dashboard visualizations
summary_all_doses <- sched_to_date %>%
  filter(Status2 %in% "Arr") %>%
  group_by(`Setting Grouper`, Site, VaxType) %>%
  summarize(Count = n()) %>%
  rename(Setting = `Setting Grouper`) %>%
  arrange(Setting, VaxType, -Count) %>%
  pivot_wider(names_from = VaxType,
              values_from = Count) %>%
  ungroup() %>%
  arrange(Setting, Site) %>%
  group_split(Setting) %>%
  # split(.$Setting) %>%
    # relocate(Setting, .after = Site) %>%
    map_df(., function(x) {
      adorn_totals(x, where = "row",
                        name = unique(x$Setting),
                        fill = "MSHS Total")})

>>>>

```{r Set directory, import reference data, and repository, include = FALSE}
# Set working directory and select raw data ----------------------------
rm(list = ls())

# Determine path for working directory
if ("Presidents" %in% list.files("J://")) {
  user_directory <- paste0("J:/Presidents/HSPI-PM/",
                           "Operations Analytics and Optimization/Projects/",
                           "System Operations/COVID Vaccine")
} else {
  user_directory <- paste0("J:/deans/Presidents/HSPI-PM/",
                           "Operations Analytics and Optimization/Projects/",
                           "System Operations/COVID Vaccine")
}

@@

# # Determine whether or not to update an existing repo
update_repo <- FALSE

# Import reference data for site and pod mappings
sched_repo_dept_mappings <- read_excel(paste0(
  user_directory,
  "/Weekly Dashboard Doses Administered",
  "/MSHS COVID-19 Vaccine Department Mappings 2022-04-11.xlsx"))

sched_repo_dept_mappings <- sched_repo_dept_mappings %>%
  select(-`Site (Historical)`, -Notes)

vaccine_admin_dept_mappings <- read_excel(paste0(
  user_directory,
  "/Hybrid Repo Sched & Admin Data",
  "/Vaccines Administered Dept Mappings 2022-07-18.xlsx"))

@@

# Vaccine types
vax_types <- c("Adult (12 yo +)", "Peds (5-11 yo)", "Peds (6 mo - 4 yo)")

# Dept grouper
dept_group <- c("Hospital POD", "MSHS Practice", "MSHS Inpatient & ED")
```

```{r Import existing repo and append as needed}
# Import schedule repository and determine last date it was updated
sched_repo_summary <- readRDS(paste0(user_directory,
                                  "/Hybrid Repo Sched & Admin Data",
                                  "/Sched Repo Summary 2020-2021 2022-07-18.RDS")
)

vax_admin_repo_summary <- readRDS(paste0(user_directory,
                                  "/Hybrid Repo Sched & Admin Data",
                                  "/Vaccines Admin Summary Jan-Jun2022 2022-07-18.RDS")
)

hybrid_repo <- rbind(sched_repo_summary,
                     vax_admin_repo_summary)

# sched_repo_update_date <- date(file.info(sched_repo_file)$ctime)

@@

  hybrid_repo <- hybrid_repo

@@

```{r Format and process schedule data}
# # Format and analyze schedule to date for dashboards ---------------------------
# sched_to_date <- sched_repo
# 
# sched_to_date <- sched_to_date %>%
#   # rename(IsPtEmployee = `Is the patient a Mount Sinai employee`) %>%
#   mutate(
#     # Update timezones to EST if imported as UTC
#     `Appt Time` = with_tz(`Appt Time`, tzone = "EST"),
#     # Determine whether appointment is for dose 1, 2, or 3 based on scheduled visit type
#     Dose = ifelse(str_detect(Type, "DOSE 1"), "Dose 1",
#                   ifelse(str_detect(Type, "DOSE 2"), "Dose 2",
#                          ifelse(str_detect(Type, "(DOSE 3)|(BOOSTER)"),
#                                 "Dose 3/Boosters", NA))),
#     # Determine appointment date, year, month, etc.
#     ApptDate = date(Date),
#     ApptYear = year(Date),
#     ApptMonth = month(Date),
#     ApptDay = day(Date),
#     ApptWeek = week(Date),
#     # Clean up department if it is duplicated in that column
#     Department = ifelse(str_detect(Department, ","),
#                         substr(Department, 1,
#                                str_locate(Department, ",") - 1),
#                         Department),
#     # Group arrived, completed, and checked out appts as arrived
#     Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
#                     "Arr", `Appt Status`),
#     # Update appointment status: classify any appts from prior days that are
#     # still in scheduled as No Shows
#     Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", 
#                      # Correct any appts erroneously marked as arrived early
#                      ifelse(ApptDate >= today & Status == "Arr", "Sch",
#                             Status)),
#     # # Determine whether the vaccine is an adult or pediatric dose
#     # VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD",
#     #                                    "1468 MADISON HOSP PEDS GEN") |
#     #                    `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
#     #                                               "DOSE 2 PEDIATRIC [1324685]") |
#     #                    Department %in% peds_school_practice_dept,
#     #                  "Peds", "Adult"),
#     # # Determine manufacturer based on immunization field and vaccine type fields
#     # Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
#     #   ifelse(Status2 != "Arr", NA,
#     #          # Any blank immunizations and any peds assumed to be Pfizer
#     #          ifelse(is.na(Immunizations) |
#     #                   VaxType %in% c("Peds"), "Pfizer",
#     #                 # Any immunizations with Moderna in text classified as Moderna
#     #                 ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
#     #                        # Any visiting docs or Johnson and Johnson
#     #                        # immunizations classified as J&J
#     #                        ifelse(str_detect(Department, "VISITING DOCS") |
#     #                                 str_detect(Immunizations,
#     #                                            "Johnson and Johnson"),
#     #                               "J&J", "Pfizer")))),
#     # Determine week number of year
#     WeekNum = format(ApptDate, "%U"),
#     # Determine appointment day of week
#     DOW = weekdays(ApptDate)
#   )
# 
# # Crosswalk sites and department grouper
# sched_to_date <- left_join(sched_to_date, dept_mappings,
#                            by = c("Department" = "Department"))
# 
# sched_to_date <- sched_to_date %>%
#   rename(Site = `New Site`)
# 
# # Determine if there are any missing departments and notify user
# unmapped_dept <- sched_to_date %>%
#   filter(is.na(Site)) %>%
#   select(Department) %>%
#   distinct()
# 
# if(nrow(unmapped_dept) > 0) {
#   stop(paste0("Check department mappings. Missing departments include ",
#               paste(as.vector(unmapped_dept$Department), collapse = ", "),
#               "."))
# }
# 
# # Determine vaccine type (adult vs peds) administered based on department, provider, and age, as appropriate
# sched_to_date <- sched_to_date %>%
#   mutate(
#     # Create a column for age grouper
#     Age_Grouper = ifelse(str_detect(Age, "\\ y.o."), "years",
#                          ifelse(str_detect(Age, "\\ m.o."), "months",
#                                 ifelse(str_detect(Age, "\\ days"), "days",
#                                        "unknown"))),
#     # Convert string with age to numeric
#     Age_Numeric = 
#       ifelse(Age_Grouper == "years",
#              as.numeric(str_extract(Age,
#                                     # pattern = "(?<=Deceased\\ \\().+(?=\\ y.o.)")),
#                                     paste("(?<=Deceased\\ \\()[0-9]+(?=\\ y.o.)",
#                                           "[0-9]+(?=\\ y.o.)",
#                                           sep = "|"))),
#              ifelse(Age_Grouper == "months",
#                     as.numeric(str_extract(Age, "[0-9]+(?=\\ m.o.)")) / 12,
#                     ifelse(Age_Grouper == "days",
#                            as.numeric(
#                              str_extract(Age, "[0-9]+(?=\\ days)")) / 365,
#                            ifelse(Age_Grouper == "unknown", 10000,
#                                   NA_integer_)))),
#     # Determine vaccine type based on department, provider/resource, and age if necessary
#     VaxType = ifelse(
#       `Vaccine Age Group` %in% "Peds" |
#         # MSBI Peds POD
#         `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
#                                    "DOSE 2 PEDIATRIC [1324685]") |
#         # Practics that administer adult and pediatric vaccines
#         (`Vaccine Age Group` %in% "Both" &
#            `Setting Grouper` %in% c("School Based Practice", "MSHS Practice") &
#            Age_Numeric < 12),
#       "Peds", "Adult"),
#     # Determine manufacturer based on immunization field and vaccine type fields
#     Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
#       ifelse(Status2 != "Arr", NA,
#              # Any blank immunizations and any peds assumed to be Pfizer
#              ifelse(is.na(Immunizations) |
#                       VaxType %in% c("Peds"), "Pfizer",
#                     # Any immunizations with Moderna in text classified as Moderna
#                     ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
#                            # Any visiting docs or Johnson and Johnson
#                            # immunizations classified as J&J
#                            ifelse(str_detect(Department, "VISITING DOCS") |
#                                     str_detect(Immunizations,
#                                                "Johnson and Johnson"),
#                                   "J&J", "Pfizer")))),
#     Site =
#       # Roll up any Moderna administered at MSB for YWCA to MSH
#       ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH",
#              # Manually classify school practices as peds or adolescent for schools administering both
#              ifelse(`Setting Grouper` %in% "School Based Practice" &
#                       VaxType %in% "Peds", "Peds School Based Practice",
#                     ifelse(`Setting Grouper` %in% "School Based Practice" &
#                              VaxType %in% "Adult",
#                            "Adolescent School Based Practice",
#                            Site))),
#     # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
#     Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c("Dose 2"), "Dose 1", Dose),
#     `Setting Grouper` = ifelse(`Setting Grouper` %in% "School Based Practice",
#                                "MSHS Practice", `Setting Grouper`))
# 
# vax_admin_start_date <- min(sched_to_date$ApptDate[
#   which(sched_to_date$Status2 %in% "Arr")])
# 
# vax_admin_end_date <- max(sched_to_date$ApptDate[
#   which(sched_to_date$Status2 %in% "Arr")])

@@

## {.tabset}
### All Patient Settings
```{r Create kable of all doses adminsitered to date}
# Summarize data for dashboard visualizations
summary_all_doses <- hybrid_repo %>%
  group_by(Setting, Site, VaxType) %>%
  summarize(Count = sum(Count, na.rm = TRUE)) %>%
  arrange(Setting, VaxType, -Count) %>%
  pivot_wider(names_from = VaxType,
              values_from = Count) %>%
  ungroup() %>%
  arrange(Setting, Site) %>%
  group_split(Setting) %>%
  # split(.$Setting) %>%
  # relocate(Setting, .after = Site) %>%
  map_df(., function(x) {
    adorn_totals(x, where = "row",
                 name = unique(x$Setting),
                 fill = "MSHS Total")})
